---
title: "23847-Lab-5"
author: '23847'
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Gen AI has not been used within the creation of this document

# Problem 1
## a)
We have $$ F_{X}(x) = \int \frac{\pi}{20} \cos{\left( \frac{(x-10)}{20}\pi \right)} dx \; = \; \sin{\left( \frac{(x-10)}{20}\pi \right)} $$ Setting $$ u = \sin{\left( \frac{(x-10)}{20}\pi \right)} \; \Rightarrow \; x = \frac{20}{\pi} \arcsin{(u)} + 10 $$ we may then generate Unif(0, 1) realizations.  

``` {r Problem 1a}
pr1 = function(n){
  return(20 / pi * asin(runif(n, 0, 1)) + 10)
}

```

## b)
We plot a histogram of the realizations generated from the pr1 function. Then the graph of the PDF is overlaid, to which we can see the sample matches the PDF. 

```{r Problem 1b}
pr1_check = function(n){
  x = pr1(n)
  hist(x , prob = TRUE, xlab = "X values", ylab = "Probability density", xlim = c(10, 20), ylim = c(0, 1))
  curve((pi / 20 * cos((x-10) / 20 * pi)), add = TRUE, col = "red")
}

pr1_check(5000)
```

## c)
To have a standard deviation below 0.002 we must first repeat the **pr1_prob** function to find a standard deviation. To calculate it we may use the **sd** function, where the number of repetitions is chosen to be 1000. Then we use a while loop to increment the number of repetitions **pr1_prob** takes until the desired deviation, here approximately 52000, is reached. 

```{r Problem 1c}
pr1_prob = function(n){
  x1 = pr1(n)
  x2 = pr1(n)
  count = sum((x1 + x2) > 25)
  return (count / n)
}

# Checking sdev 
nreps = 10000
std_dev = 1
while(std_dev > 0.002){
  std_dev = sd(replicate(1000, pr1_prob(nreps)))
  nreps = nreps + 1000
}
print(nreps)

pr1_prob(52000) # Estimate 

```

# Problem 2
## a)
We calculate the times $$ X_{i} $$ by comparing the values of $$ \sum_{1}^{i} T_{i} \stackrel{?}{\geq} 15i $$, if it is greater then the appointment overran and the following appointments will be delayed. 
Then we iterate over the values in Y := X + T, subtracting their arrival times to calculate W, which is then directly inputted into the array. To check the function was working as intended a test with T = (16, 15, 30) was inputted, resulting in W = (16, 16, 32) to be returned, as desired.  

``` {r Problem 2a}
gentimes = function(n){  
  z = array(0, c(n, 3))
  for(i in 1:n){
      T = runif(3, 5, 30) 
      X = c(0, max(T[1], 15), max(T[1] + T[2], 30) )
      z[i, ] = (X + T)[1:3] - 15 * (0:2) 
  }
  return (z)
}
```

## b)
The plot has a linear correlation, which can be seen by adding a line of best fit. This implies the variables are not independent.

``` {r Problem 2b}
times = gentimes(2000)
plot(times[1:2000, 1], times[1:2000, 3], xlab = "W1", ylab = "W3")
abline(lm(times[1:2000, 3] ~ times[1:2000, 1]), col = "red")
```

## c)
Firstly we calculate the expectaions as given in the problem sheet. Then we calculate the variance using $$ Var(W_{i}) = E[W_{i}^{2}] - E[W_{i}]^{2} $$ and then the covariance, $$ Cov(W_{1},W_{3}) = E[W_{1}W_{3}] - E[W_{1}]E[W_{3}] $$. Using these two values we may then calculate the correlation coefficient, $$ Corr(W_{1},W_{3}) = \frac{Cov(W_{1},W_{3})}{\sqrt{Var(W_{1})Var(W_{3})}} $$ This gives us a correlation coefficient of approximately 0.48. 

``` {r Problem 2c}
n = 20000
wPairs = gentimes(n)[, -2]
E_W1 = 1/n * sum(wPairs[ , 1])
E_W1_2 = 1/n * sum(wPairs[ , 1]^2)

E_W3 = 1/n * sum(wPairs[ , 2])
E_W3_2 = 1/n * sum(wPairs[ , 2]^2)

E_W1_W3 = 1/n * sum(wPairs[ , 1] * wPairs[ , 2])

var_W1 = E_W1_2 - E_W1^2
var_W3 = E_W3_2 - E_W3^2
covar_W1_W3 = E_W1_W3 - (E_W1 * E_W3)

corr = covar_W1_W3 / sqrt(var_W1 * var_W3)
print(corr)
```





