---
title: "Lab 4 exercises"
output: word_document
date: "2024-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Gen AI has not been used within the creation of this document

# Problem 1
## a)
First we calculate the first moment using the formula $$ \mu = \mathbb{E}[X] = \frac{1}{n}\sum_{i=1}^{n}x_{i} $$ then we calculate the second moment using the formula $$ \sigma^{2} + \mu^{2}  = \mathbb{E}[X^{2}] = \frac{1}{n}\sum_{i=1}^{n}x_{i}^{2} $$ where x_i represent the elements of the npsold vector. We can then rearrange to get sigma^2, denoted as **sdev**.
The data does not appear to follow the distribution however this could be expected as the standard deviation is quite large. 
```{r npSold, include=FALSE}
npsold =
 c(158, 196, 182, 190, 179, 162, 153, 179, 190, 194,
   229, 204, 175, 211, 188, 163, 224, 150, 187, 155,
   200, 185, 173, 202, 172, 184, 190, 189, 175, 190,
   175, 190, 154, 162, 176, 202, 190, 211, 188, 211,
   209, 189, 168, 178, 176, 188, 157, 188, 178, 161,
   179, 162, 188, 187, 195, 215, 162, 171, 184, 208,
   221, 190, 166, 159, 198, 163, 174, 183, 186, 180,
   159, 213, 163, 210, 205, 157, 225, 211, 166, 155,
   147, 175, 182, 215, 230, 181, 203, 184, 206, 178,
   221, 173, 183, 175, 146, 199, 182, 154, 181, 223)
```

```{r Problem 1a}
# npsold not included to keep the code clean
m1 = sum(npsold)/length(npsold)
m2 = sum(npsold^2)/length(npsold)
sdev = sqrt(m2 - m1^2)
hist(npsold, breaks = 25, xlab = "Frequency", ylab = "Frequency density", main = "Histogram of Newspapers sold", prob = TRUE)
curve(dnorm(x, m1, sdev), add = TRUE, col = "red")
```

## b)
When making a Q-Q plot you must sort the observed values and compare these to $$ F^{-1}(frac{i}{n+1}) $$. When the data is a good fit it will be observed to be close to the line y  = x. 
```{r Problem 1b}
x= sort(npsold)
i = c(1:100)
y = qnorm(i/101, m1, sdev)
plot(x,y)
curve(x^1, add = TRUE, col = "red")


obs1 = sort(rnorm(100, m1, sdev))
plot(obs1, y)
curve(x^1, add = TRUE, col = "red")
obs2 = sort(rnorm(100, m1, sdev))
plot(obs2, y)
curve(x^1, add = TRUE, col = "red")
obs3 = sort(rnorm(100, m1, sdev))
plot(obs3, y)
curve(x^1, add = TRUE, col = "red")
```

## c)
For the case n>200 we may simplify the return, expanding from 200 * 0.3 + (n-200) * 0.1 to 40 + n * 0.1. 

```{r Problem 1c}
# i
cost = function(n){
  if((0 <= n) && (n <= 200)){
    return(n * 0.3)
  } 
  else if(n > 200){
    return(40 + n * 0.1)
  }
}

# ii
profit = function(n, d){
  return (d - cost(n))
}

# iii
averageProfit = function(n, mu, sigma, nreps){
  return (sum(rnorm(nreps, mu, sigma)) / n)  
}
```

## d)
To find the optimal number of papers we may use mu and sigma from **a)**, then entering , to see which value has the highest average profit. 

```{r Problem 1d}
avgProf = c(averageProfit(seq(from = 5, to = 100, by = 5), m1, sdev, 100))
```
