---
title: "Lab 4 exercises"
output: word_document
date: "2024-03-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Gen AI has not been used within the creation of this document

# Problem 1
## a)
First we calculate the first moment using the formula $$ \mu = \mathbb{E}[X] = \frac{1}{n}\sum_{i=1}^{n}x_{i} $$ then we calculate the second moment using the formula $$ \sigma^{2} + \mu^{2}  = \mathbb{E}[X^{2}] = \frac{1}{n}\sum_{i=1}^{n}x_{i}^{2} $$ where x_i represent the elements of the npsold vector. We can then rearrange to get $$ \sigma^{2} $$ , denoted as **sdev**.
The data does not appear to follow the normal distribution however, this could be expected as the standard deviation is quite large. 
```{r npSold, include=FALSE}
npsold =
 c(158, 196, 182, 190, 179, 162, 153, 179, 190, 194,
   229, 204, 175, 211, 188, 163, 224, 150, 187, 155,
   200, 185, 173, 202, 172, 184, 190, 189, 175, 190,
   175, 190, 154, 162, 176, 202, 190, 211, 188, 211,
   209, 189, 168, 178, 176, 188, 157, 188, 178, 161,
   179, 162, 188, 187, 195, 215, 162, 171, 184, 208,
   221, 190, 166, 159, 198, 163, 174, 183, 186, 180,
   159, 213, 163, 210, 205, 157, 225, 211, 166, 155,
   147, 175, 182, 215, 230, 181, 203, 184, 206, 178,
   221, 173, 183, 175, 146, 199, 182, 154, 181, 223)
```

```{r Problem 1a}
# npsold not included to keep the code clean
m1 = sum(npsold)/length(npsold)
m2 = sum(npsold^2)/length(npsold)
sdev = sqrt(m2 - m1^2)
hist(npsold, breaks = 40, xlab = "Frequency", ylab = "Frequency density", main = "Histogram of Newspapers sold", prob = TRUE)
curve(dnorm(x, m1, sdev), add = TRUE, col = "red")
```

## b)
When making a Q-Q plot you must sort the observed values and compare these to $$ F^{-1}(frac{i}{n+1}) $$. When the data is a good fit it will be observed to be close to the line y  = x. 
```{r Problem 1b}
x= sort(npsold)
i = c(1:100)
y = qnorm(i/101, m1, sdev)
plot(x,y)
curve(x^1, add = TRUE, col = "red")


obs1 = sort(rnorm(100, m1, sdev))
plot(obs1, y)
curve(x^1, add = TRUE, col = "red")
obs2 = sort(rnorm(100, m1, sdev))
plot(obs2, y)
curve(x^1, add = TRUE, col = "red")
obs3 = sort(rnorm(100, m1, sdev))
plot(obs3, y)
curve(x^1, add = TRUE, col = "red")
```

## c)
### i.
For the case n>200 we may simplify the return, expanding from 200 * 0.3 + (n-200) * 0.1 to 40 + n * 0.1. 
### ii.

### iii.
For the average profit we will calculate the mean value by summing the outputs of the profit vector and dividing it by the number of reps 

```{r Problem 1c}
# i
cost = function(n){
  r = numeric(length(n))
  for(i in 1:length(n)){
      if((0 <= n[i]) && (n[i] <= 200)){
        r[i] = 0.3 * n[i]
      } 
      else if(n[i] > 200){
        r[i] = 40 + n[i] * 0.1
      }
  }
  return (r)
}

cost = function(n){
  if((0 <= n) && (n <= 200)){
    return(n * 0.3)
  } 
  else if(n > 200){
    return(40 + n * 0.1)
  }
}


# ii
profit = function(n, d){
  return (d - cost(n))
}

# iii
averageProfit = function(n, mu, sigma, nreps){
  s = 0
  for(i in 1:length(n)){
    s = s + profit(n[i], rnorm(nreps, mu, sigma))
  }
  return (s / nreps)  
}
```

## d)
To find the optimal number of papers we may use mu and sigma from **a)**, then entering , to see which value has the highest average profit. 

```{r Problem 1d}
avgProf = c(averageProfit(seq(from = 5, to = 100, by = 5), m1, sdev, 100))
```

## e)
Relevant?
We have $$ averageProfit := A = d_{1} + ... + d_{n} \, with \, d ~ N(\mu, \sigma^{2}) $$ Using the addition formula for normally distributed RVs we get $$ A ~ N(1*\mu + ... + 1*\mu, 1^{2}\sigma^{2} + ... + 1^{2}\sigma^{2}) = N(n*\mu, n*sigma^{2}) $$ This distribution represents the average profit for the purchase of n papers. 

We would need to know the time it takes Erik to sell his papers, whether it is linear, so each paper takes the same amount of time, or if the time follows its own distribution. Once this information is obtained we could set up a relation between profits and sell time, a relation we would then be able to minimise the time for a given profit. 

