---
title: "Lab 2 exercises"
output: html_document
date: "2024-03-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
------------------------------------------------------------------------------------------------------------------------------
Gen AI has not been used in this assignment

## Exercise 1
a) we use the function rnorm(n, m, s) to generate 1000 realizations

b) the hist function is then used to plot the realizations vector with prob = True to show density instead of frequency

c) the curve function is used to add the PDF of the normal distribution N(3, 4^2) to the histogram.  

``` {r Exercise 1}
x = rnorm(1000, 3, 4)
hist(x, prob = TRUE, main = "Observed X values", xlab = "X values", ylab = "Density of the distribution")
curve(1/(4 * sqrt(2 * pi)) * exp(-(x - 3)^2 / 32), add = TRUE, col = "red")

```
------------------------------------------------------------------------------------------------------------------------------
## Exercise 2
### a) 
number of trials, n := 1000
probability of rolling a 4, p := 1/6 
The probability P(150 <= X <= 180) = P(X <= 180) - P(X < 150)
                                      = P(X <= 180) - P(X <= 149)
                                      = pbinom(180, n, p) - pbinom(149, n, p)
   
   
### b)
We have N(np, np(1 - p)), without continuity correction we can reuse the code from part a,
changing the function from pbinom to pnorm. For the continuity correction we may use 
P(X < 150) ≈ P(Y ≤ 150 − 0.5) instead of correcting to P(X <= 149). 
   
``` {r Exercise 2}
n = 1000
p = 1/6
p1 = pbinom(180, n, p) - pbinom(149, n, p)
nCont = pnorm(180, n*p, n*p*(1 - p)) - pnorm(149, n*p, n*p*(1 - p))
Cont = pnorm(180 + 0.5, n*p, n*p*(1 - p)) - pnorm(150 - 0.5, n*p, n*p*(1 - p))
```
------------------------------------------------------------------------------------------------------------------------------
## Exercise 3
### a) 
m := 2^32
a := 1664525
c := 1013904223
We have a recursive sequence so we shall use a for loop and take the value
of the previous index. To make it less computationally expensive we will 
allocate the memory of X beforehand so that a copy needn't be created each iteration
This can be done with the numeric(x) function. I have chosen the seed as X_1 = 1 
U is then defined to be the set of X / m which can then be plotted with the hist function.

### b) 
Once again we allocate the memory for the vectors beforehand. 
Next we want to choose replicable random numbers from the set U. Here i have chosen to take values
every multiple of 3 for the X coordinates and every multiple of 5 for the Y coordinates. 
Any sequence can be used here as long as it produces values within [1, 10000]
   
### c) 
The points tend to form more clusetered patches in some areas with large spaces in other areas
   
   
### d) 
Randomly created sequences chosen, a +1 was required on a few as a_1 would evaluate to zero   
\begin{table}[h]
\centering
\begin{tabular}{|c|c|c|}
\hline
Plot \# & X sequence & Y sequence \\
\hline
1 & \(4n + 1\) & \(10001 - n^2\) \\
2 & \(\text{floor}(10 \times \log_{10}(n)) + n\) & \(\text{ceil}(2^{(n/9)})\) \\
3 & \(\text{floor}(\exp(n/11))\) & \(\frac{1}{2}n(n-1) + 1\) \\
4 & \(1 + n + n^2\) & \(10n\) \\
\hline
\end{tabular}
\caption{Description of Sequences}
\label{tab:sequences}
\end{table}
``` {r Exercise 3}
#a)
m = 2^32
a = 1664525
c = 1013904223
X = numeric(10000)
X[1] = 1
for(i in 2:10000){
  X[i] = ((a * X[i - 1] + c) %% m)
}
U = c(X / m)
hist(U, prob = TRUE, main = "Linear congruential generator")

#b)
Xcoord = numeric(100)
Ycoord = numeric(100)
for(i in 1:100){
  Xcoord[i] = U[3 * i] * 10
  Ycoord[i] = U[5 * i] * 10
}
plot(Xcoord, Ycoord, main = "Acorn map")

#d)
par(mfrow=c(2,2)) 
Xcoord = numeric(100)
Ycoord = numeric(100)
# Plot 1
for(i in 1:100){
  Xcoord[i] = U[4 * i + 1] * 10
  Ycoord[i] = U[10001 - i^2] * 10
}
plot(Xcoord, Ycoord, main = "Acorn map")

# Plot 2
for(i in 1:100){
  Xcoord[i] = U[floor(10 * log10(i)) + i] * 10
  Ycoord[i] = U[ceiling(2^(i / 9))] * 10
}
plot(Xcoord, Ycoord, main = "Acorn map")

#Plot 3
for(i in 1:100){
  Xcoord[i] = U[floor(exp(i / 11))] * 10
  Ycoord[i] = U[1/2*i*(i-1) + 1] * 10
}
plot(Xcoord, Ycoord, main = "Acorn map")

#Plot 4
#Plot 3
for(i in 1:100){
  Xcoord[i] = U[1 + i + i^2] * 10
  Ycoord[i] = U[10 * i] * 10
}
plot(Xcoord, Ycoord, main = "Acorn map")
```